---
title: "Ratio Estimators Analysis"
format: pdf
author: "Chenming Zhao"
---
  
## 1. Data Acquisition
The dataset used in this analysis is available [https://usa.ipums.org/usa/index.shtml]. We use IPUMS to access the 2022 ACS and focus on each state (STATEICP) that had a doctoral degree as their highest educational attainment (EDUC).We specified state-level data by selecting "HOUSEHOLD" > "GEOGRAPHIC" and added "STATEICP" to our cart. For individual-level data, we went to "PERSON" and added "EDUC" and "SEX" to our cart. We then downloaded and saved it locally (e.g., "usa_00001.csv.gz") for use in R.

## 2. Overview of Ratio Estimators
Ratio estimators provide a method for estimating population parameters by using auxiliary information. 

```{r, message=FALSE}
#| echo: false
# Load the necessary libraries
library(dplyr)
library(haven)
library(tidyverse)
library(labelled)

# Read the dataset (assuming it's a CSV)
data <- read.csv("/Users/dragon/Desktop/usa_00001.csv.gz")
```
The number of respondents were there in each state (STATEICP) that had a doctoral degree as their highest educational attainment (EDUC)
```{r, message=FALSE}
#| echo: false

cleaned <- 
  data |>
  select(STATEICP,EDUC, EDUCD) |>
  to_factor()

doctoral_counts <- cleaned |>
  filter(EDUCD == 116) |>  # Filter rows with "Doctoral degree"
  group_by(STATEICP) |>
  summarise(doctoral_count = n()) |>  # Count respondents with a doctoral degree in each state
  ungroup()
doctoral_counts
```

```{r}
#| echo: false
state_41_data <- data %>%
  filter(STATEICP == 41)

total_respondents_state_41 <- state_41_data %>%
  summarise(Total_Respondents = n())

#Calculate the number of respondents with EDUC == 6 (doctoral degree) in state 41
educ_respondents_state_41 <- state_41_data %>%
  filter(EDUC == 6) %>%
  summarise(educ_Respondents = n())

#Calculate the ratio of doctoral respondents to total respondents in state 41
ratio_state_41 <- educ_respondents_state_41$educ_Respondents / total_respondents_state_41$Total_Respondents

#Apply the ratio to estimate the total number of respondents with EDUC == 6 in each state
# Group by state and calculate the number of doctoral respondents (EDUC == 6) in each state
statewise_educ_respondents <- data %>%
  filter(EDUC == 6) %>%
  group_by(STATEICP) %>%
  summarise(educ_Respondents = n())

#Estimate the total number of respondents with a doctoral degree in each state using the ratio from state 41
statewise_estimates <- statewise_educ_respondents %>%
  mutate(Estimated_Total_Respondents = educ_Respondents / ratio_state_41)

#View the results
head(statewise_estimates)

```


## 3. Estimates and Actual Number of Respondents
Here are the estimates and the actual numbers based on the analysis:

```{r}
#| echo: false
actual_state_Respondents <- data %>% 
  group_by(STATEICP) %>%
  summarise(Actual_Total_responses = n())

head(actual_state_Respondents)
```

```{r}
#| echo: false
comparison_table <- merge(statewise_estimates %>% select(STATEICP, Estimated_Total_Respondents),
                          actual_state_Respondents, by = "STATEICP")

#View the comparison table (showing both estimated and actual respondents)
head(comparison_table)

```

## 4. Explanation of Differences
The differences between the estimates and the actual numbers could be due to several factors, including:
1. We assumed the relationship between the number of respondents with EDUC = 6 degrees and the total number of respondents in one state (in this case, state 41) applies similarly to all other states, but this could vary by different culture and education background.
2. Each state has its own economic and cultural factors that affect education levels.
3. The dataset from state 41 may not represent the broader national trend due to sampling bias. If the respondents in state 41 are not reflective of the national population, the ratio derived from this state will introduce bias when applied to other states.
4. Ratio estimators rely on a linear relationship between the number of doctoral degree holders and the total population, which may not be valid across all states. In reality, the relationship may be more complex, with certain factors (e.g., urbanization, economic development) influencing education levels in nonlinear ways.

quarto render document.qmd --to pdf
